---
title: Ergodicity of the CaPso model
author: Mario Martínez Molina
output:
    html_document:
        theme: united
        highlight: zenburn
        df_print: paged
---

## Packages
```{r}
library(rcapso)
library(ggplot2)
library(dplyr)
```

```{r}
datasets <- Sys.getenv(c("DATASETS"))
```

## Introduction

Orbits belonging to a chaotic dynamical system are highly unstable (in a
chaotic dynamical system orbits diverge exponentially). However, from a
statistical perspective some form of predictability can emerge. To illustrate
this idea we can divide the $[0, 1]$ interval in $N$ subintervals, and
calculate the number of points in an orbit of the CaPso model in each interval.
For such purpose, I've selected the parameters of Figure 6.9 which result in an
orbit similar to a limit cycle, albeit somewhat unstable. Note that I'm
removing the first 1000 points to get rid off any transient.

```{r}
mf_ez_2 <- rcapso_compute_mf(use_reg = FALSE, num_iter = 10000,
                             psi0 = 1, phi0 = 0.001, alpha = 0.1,
                             ey = 1, ez = 2) %>%
slice_tail(n = 9000)
head(mf_ez_2)
```

```{r}
filename <- paste(datasets[1], "capso/code_test/pcg_fig_6_9_a_0_0.csv",
                  sep = "\\")
lattice_size <- 262144

sm_ez_2 <- read.csv(filename) %>%
    slice_tail(n = 9000) %>%
    transmute(preys_norm = Preys / lattice_size,
              preds_norm = Predators / lattice_size)
head(sm_ez_2)
```

```{r, dev = "svg"}
ggplot(data = mf_ez_2) +
    geom_path(mapping = aes(x = Preys, y = Predators)) +
    theme(aspect.ratio = 1)
```

```{r, dev = "svg"}
ggplot(data = sm_ez_2) +
    geom_path(mapping = aes(x = preys_norm, y = preds_norm)) +
    theme(aspect.ratio = 1)
```

If the orbit were completely random, then an histogram of the points in the
orbit, as described before, would approximate a uniform distribution. Instead,
we can see that in an orbit of the CaPso model some points are more likely to
appear than others.

```{r, dev = "svg"}
ggplot(data = mf_ez_2, aes(x = Preys)) +
    geom_histogram(aes(y = after_stat(density)), binwidth = 0.005) +
    geom_density(linewidth = 1) +
    theme(aspect.ratio = 9 / 16)
```

```{r, dev = "svg"}
ggplot(data = sm_ez_2, aes(x = preys_norm)) +
    geom_histogram(aes(y = after_stat(density)), binwidth = 0.005) +
    geom_density(linewidth = 1) +
    theme(aspect.ratio = 9 / 16)
```

For the mean field model the orbits are restricted to the interval $[0.2218707,
0.6561584]$, meanwhile simulations are restricted to the interval $[0.3237381,
0.6927452]$. Nevertheless, it appears that the singularities of the orbit are
dense in this interval.

```{r}
columns <- c("Preys", "Predators")
sample_rows <- c(190:200)

rm(ergdf)
for (file in list.files(c(paste(datasets[1],
                                "capso/ergodicity2",
                                sep = "\\")))) {
    filepath <- paste(datasets[1], "capso/ergodicity2", file, sep = "\\")

    if (!exists("ergdf")) {
        ergdf <- read.csv(filepath, header = FALSE,
                          col.names = columns)
        ergdf <- ergdf[sample_rows, ]
    } else {
        tmp   <- read.csv(filepath, header = FALSE,
                          col.names = columns)
        ergdf <- rbind(ergdf, tmp[sample_rows, ])
        rm(tmp)
    }
}

ergdf
```

```{r}
ergnorm <- ergdf %>%
    transmute(preys_norm = as.numeric(Preys) / lattice_size,
              preds_norm = as.numeric(Predators) / lattice_size)
ergnorm
```

```{r}
hist(ergnorm$preys_norm, freq = FALSE, xlab = "x(t)", ylab = "density",
     br = seq(0, 1, by = 0.005))
```

```{r, dev = "svg"}
test <- read.csv(paste(datasets[1], "capso/ergodicity2", "density_0.492000.csv",
                       sep = "\\"))

test <- test %>%
    slice_tail(n = 100) %>%
    transmute(preys_norm = Preys / lattice_size,
              preds_norm = Predators / lattice_size)

test
```

```{r}
ggplot(data = test) +
    geom_point(mapping = aes(x = preys_norm, y = preds_norm)) +
    theme(aspect.ratio = 1)
```
